Considering scale in multilayer animal social networks

### 2020-08-17 paper/manuscript.Rmd
minor edits

# Introduction
Our inference regarding the processes underlying ecological patterns
are directly linked to the scale at which they are observed [@Levin_1992;
@Allen_2015]. Animals are influenced across scales through ecological processes such as
trophic interactions, resource availability
[@Legendre_1993; @Chave_2013], and global climate and productivity [@Field_2009].
Multiscale ecology has been integrated into species distribution modelling
[@Elith_2009], habitat selection [@Mayor_2007], and food webs [@Sugihara_1989].
We posit that it is important to consider the scale dependence of 
animal behaviour within the social environment to effectively understand 
complex social systems.

@@

We examined scale-dependent multilayer networks motivated by behavioural,
landscape, and spatial ecology to and applied this framework to caribou
(*Rangifer tarandus*). First, we test to
what degree the social scale of relationships, spatial scale of landscapes, and
temporal scale of analysis influences our ability to interpret complex social
systems. Second, we address how variation in the social, spatial, and temporal
scale improves our understanding of the relationship between social association
and seasonal resource availability. We conclude with potential future
developments and applications of scale in multilayer networks.

@@

## Caribou socioecology
Caribou are gregarious ungulates with fission-fusion dynamics [@Lesmerises_2018]
that display temporal and spatial variation in social networks [@Peignier_2019].
In winter, caribou dig holes, termed craters, to access forage beneath the snow
[@Bergerud_1974]. Cratering is presumed to be a costly behaviour, and as a
result, caribou tend to occupy and re-use craters once they are established.
Importantly, craters exist on the landscape at multiple scales: the crater scale
(a single crater), the feeding area scale (multiple craters in close proximity),
and the winter range scale (all craters within an individual’s range)
[@Mayor_2009]. Access to forage is therefore heterogeneous in winter and depends
on snow cover and distribution of craters on the landscape.

@@

## Landscape data and habitat classification
Land cover data were provided by the Newfoundland and Labrador Wildlife Division
[@Integrated_2013]<!--TODO: check ref-->. Available land cover classification
included nine habitat types at 30 m resolution. We reclassified the land cover
types into three categories using the R package `raster` [@Hijmans_2019]:
foraging, forest, and open habitats. Foraging habitat consisted of lichen,
forest habitat consisted of coniferous forest, conifer scrub, broad leaf forest,
and mixed-wood forest, while open habitat consisted of wetland, rocky barrens,
and anthropogenic. Water habitat was excluded from all subsequent analyses.

@@

We generated multilayer networks across a series of social distance thresholds
for group assignment. Social distance thresholds (5, 25, 50, 75, 100, 250 and
500 m) represent a range of sensory modalities of caribou from visual 
to olfactory and auditory. Multilayer networks consisted
of the social association between `r DT[, uniqueN(get(idcol))]` individuals
across three habitat layers (foraging, forest, and open) for the entire study
period. At the finest scale, individuals within 5 m of one another were
considered in the same group, whereas at the coarsest scale, individuals within
500 m of one another were considered in the same group. Increasing social
distance threshold is a proxy for potential behavioural interactions that occur
across these distances. For every social distance threshold, we calculated
individual graph strength within each habitat layer, and edge overlap and
multidegree across habitat layers.

@@

Habitat selection, the use of available resources or landscape features by
animals [@Manly_1993; @Boyce_1999], is a scale-dependent process [@Mayor_2009]
that we can observe in multilayer networks.
Social structure and habitat selection are intimately related and individuals
must share space to interact [@Webber_2018]. We found edge overlap was stable in
open and forest layers but decreased in foraging layers as spatial
scale increased (Figure \@ref(fig:lcres) D). Graph strength had greater individual
variation in open and foraging layers than in forest layers (Figure
\@ref(fig:lcres) E). Furthermore, we observed a decrease in the proportion of relocations in
foraging habitat with increasing spatial resolution 
(`r count[res == 30 & lc == 3, percent]`% at 30 m - `r count[res == 1000 & lc == 3, percent]`% at 1000
m). As the foraging layer had similar edge overlap and graph strength as the
open layer at 30 m spatial resolution, spatial scale therefore played a
key role in our multilayer social networks. 
Johnson [-@Johnson_1980] identified four orders, or scales,  of habitat selection: (1) the
geographical range of a species, (2) the home range of an individual, the (3)
habitat patches within the home range and (4) specific resources within a
habitat patch. Scales of habitat selection could be relevant for multilayer networks and our
integration of habitat within a multilayer context could be an important first
step. Specifically, if individuals select habitats
at the home range scale - e.g. at coarser resolution, it may be relevant to assess social structure using
home range overlap. By contrast, in cases where individuals select habitats at
the patch-scale, fine-scale measures of sociality may be more relevant. Moving
past this hierarchical perspective, Mayor et al. [-@Mayor_2007] emphasized the
importance of considering a broad continuum of scales to reveal scale-dependent
selection and avoid the bias of predefined scales assumed to associate with
certain behaviours. We highlight the concept of scale of scale in multilayer 
networks should be approached based on explicit and precise definitions
to ensure robust measurements.

@@

Recognition of scale as a fundamental concept in ecology [@Levin_1992] has lead
to an increased attention to the scale at which ecological processes are
observed [@Schneider_2001; @Chave_2013]. Recent technological advances in
remotely sensed imagery [@Drusch_2012] and animal tracking devices
[@Borger_2020; @B_rger_2016; @Wikelski_2007] have increased the availability of
complex, high-resolution data. We highlight ways in which scale should be
considered when building social networks from both observed and remotely sensed
data. Our multilayer network analysis of caribou sociality incorporating high
resolution GPS telemetry data and spatial land cover data is grounded in the
context of observed animal behaviour as the scales we chose to explore, e.g.,
social, spatial, and temporal, are scales that have
also been investigated using traditional behavioural ethological methods.

Multilayer networks remain a nascent but powerful tool in animal behavioural
ecology. Silk et al. [-@Silk_2018] summarized some key multilayer questions in
animal behaviour research and, based on our study of scale, 
we conclude by posing an additional set of questions on
the future of multilayer networks through the lens of scale and behavioural,
landscape, and spatial ecology:

>>>>

# Introduction
Our inferences regarding the processes underlying ecological patterns
are directly linked to the scale at which they are observed [@Levin_1992;
@Allen_2015]. Animals are influenced across scales by ecological processes such as
trophic interactions, resource availability
[@Legendre_1993; @Chave_2013], and global climate and productivity [@Field_2009].
Multiscale ecology has been integrated into species distribution modelling
[@Elith_2009], habitat selection [@Mayor_2007], and food webs [@Sugihara_1989].
We posit that it is important to consider the scale dependence of 
animal behaviour within the social environment to effectively understand 
complex social systems.

@@

We examined scale-dependent multilayer networks of behavioural,
landscape, and spatial ecology and applied this framework to caribou
(*Rangifer tarandus*). First, we test to
what degree the social scale of relationships, spatial scale of landscapes, and
temporal scale of analysis influences our ability to interpret complex social
systems. Second, we address how variation in the social, spatial, and temporal
scale improves our understanding of the relationship between social association
and seasonal resource availability. We conclude with potential future
developments and applications of scale in multilayer networks.

@@

## Caribou socioecology
Caribou are gregarious ungulates with fission-fusion dynamics [@Lesmerises_2018]
that display temporal and spatial variation in social networks [@Peignier_2019].
Caribou dig holes, termed craters, in winter to access forage beneath the snow
[@Bergerud_1974]. Cratering is presumed to be a costly behaviour, and as a
result, caribou tend to occupy and re-use craters once they are established.
Importantly, craters exist on the landscape at multiple scales: the crater scale
(a single crater), the feeding area scale (multiple craters in close proximity),
and the winter range scale (all craters within an individual’s range)
[@Mayor_2009]. Access to forage is therefore heterogeneous in winter and depends
on snow cover and distribution of craters on the landscape.

@@

## Landscape data and habitat classification
Land cover data were provided by the Newfoundland and Labrador Wildlife Division
[@Integrated_2013]<!--TODO: check ref-->. Available land cover classification
included nine habitat types at 30 m resolution. We reclassified the land cover
types into three categories using the R package `raster` [@Hijmans_2019]:
foraging, forest, and open habitats. Foraging habitat consisted of lichen,
forest habitat consisted of coniferous forest, conifer scrub, broad leaf forest,
and mixed-wood forest, while open habitat consisted of wetland, rocky barrens,
and anthropogenic areas. Water was excluded from all subsequent analyses.

@@

We generated multilayer networks across a series of social distance thresholds
for group assignment. Social distance thresholds (5, 25, 50, 75, 100, 250 and
500 m) represent a range of sensory modalities of caribou from visual 
to olfactory and auditory. Multilayer networks consisted
of the social association between `r DT[, uniqueN(get(idcol))]` individuals
across three habitat layers (foraging, forest, and open) for the entire study
period. At the finest scale, individuals within 5 m of one another were
considered in the same group, whereas at the coarsest scale, individuals within
500 m of one another were considered in the same group. Increasing social
distance threshold is a proxy for potential behavioural interactions (eg. vocal communication)
that occur
across these distances. For every social distance threshold, we calculated
individual graph strength within each habitat layer, and edge overlap and
multidegree across habitat layers.

@@

Habitat selection, the use of available resources or landscape features by
animals [@Manly_1993; @Boyce_1999], is a scale-dependent process [@Mayor_2009]
that we can observe in multilayer networks.
Social structure and habitat selection are intimately related and individuals
must share space to interact [@Webber_2018]. We found edge overlap was stable in
open and forest layers but decreased in foraging layers as spatial
scale increased (Figure \@ref(fig:lcres) D). Graph strength had greater individual
variation in open and foraging layers than in forest layers (Figure
\@ref(fig:lcres) E). Furthermore, we observed a decrease in the proportion of relocations in
foraging habitat with increasing spatial resolution 
(`r count[res == 30 & lc == 3, percent]`% at 30 m - `r count[res == 1000 & lc == 3, percent]`% at 1000
m). As the foraging layer had similar edge overlap and graph strength as the
open layer at 30 m spatial resolution, spatial scale therefore played a
key role in our multilayer social networks. 
Johnson [-@Johnson_1980] identified four orders, or scales,  of habitat selection: (1) the
geographical range of a species, (2) the home range of an individual, the (3)
habitat patches within the home range and (4) specific resources within a
habitat patch. Scales of habitat selection could be relevant for multilayer networks and our
integration of habitat within a multilayer context could be an important first
step. Specifically, if individuals select habitats
at the home range scale - e.g. at coarser resolution, it may be relevant to assess social structure using
home range overlap. By contrast, in cases where individuals select habitats at
the patch-scale, fine-scale measures of sociality may be more relevant. Moving
past this hierarchical perspective, Mayor et al. [-@Mayor_2007] emphasized the
importance of considering a broad continuum of scales to reveal scale-dependent
selection and avoid the bias of predefined scales assumed to associate with
certain behaviours. We highlight the concept of scale in multilayer 
networks should be approached based on explicit and precise definitions
to ensure robust measurements.

@@

<!-- rm:
Recognition of scale as a fundamental concept in ecology [@Levin_1992] has lead
to an increased attention to the scale at which ecological processes are
observed [@Schneider_2001; @Chave_2013]. Recent technological advances in
remotely sensed imagery [@Drusch_2012] and animal tracking devices
[@Borger_2020; @B_rger_2016; @Wikelski_2007] have increased the availability of
complex, high-resolution data. We highlight ways in which scale should be
considered when building social networks from both observed and remotely sensed
data. Our multilayer network analysis of caribou sociality incorporating high
resolution GPS telemetry data and spatial land cover data is grounded in the
context of observed animal behaviour as the scales we chose to explore, e.g.,
social, spatial, and temporal, are scales that have
also been investigated using traditional behavioural ethological methods. 
-->

Multilayer networks remain a nascent but powerful tool in animal behavioural
ecology. Silk et al. [-@Silk_2018] summarized some key multilayer questions in
animal behaviour research and, based on our study of scale, 
we conclude by posing an additional set of questions centered on scale in
behavioural, landscape, and spatial ecology:
